---
title: "Bookbinders Book Club"
author: "Jordi López Sintas"
date: "3/19/2018"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# Los datos

The case Bookbinder Book Club, BBBClub, illustrates the use of three ways that can be use to select the potential clients that will be the target of our marketing efforts: (1) the traditional RFM model, (2) the linear model, and (3) the logistic model.

We need two data sets, one historic data set with the past behavior of a sample of clients, and another data set with the data of possible new clients of a book, in this case.

## The historic data set with the past behavior of a sample of clients

```{r data}

BBBClub.choice<-read.delim(
  "R-Bookbinders Book Club Data (Customer Choice).txt")
BBBClub.holdout<-read.delim(
  "R-Bookbinders Book Club Data (Customer Choice) Holdout Sample.txt")

knitr::kable(
  t(head(BBBClub.choice)), 
  digits=2, 
  caption = "Descripción de las 6 primeras empresas de los datos históricos")

```

## The data set with prospects

```{r}
knitr::kable(
  t(head(BBBClub.holdout)), 
  digits=2, 
  caption = "Descripción de las 6 primeras empresas de los nuevos datos ")
```


# RFM model

The RFM is based on the Recency, Frequency and Monetary value of old clients with similar characteristics.

- Recency: Tiempo transcurrido desde la última compra (< 2 wks, 2-6 wks, 7-8 wks, and  > 12 wks)
- Frequency: Número de compras realizadas desde una determinada fecha hasta la actualidad (10, 8, 6, 3, 1)
- Monetary Value: Valor de las compras desde una determina fecha (> 500, 300-500, 200-300, 100-200, < 200)

## First we apply the RFM model to the new data (prospects)

Asignar una puntuación a R, F y M según la experiencia. Por ejemplo:

*Recency*:

últimos 3meses, 25 puntos
3-6 meses, 20 puntos
6-9 meses, 10 putnos
10-18 meses, 5, puntos
Más de 18 meses, 0 puntos

Aplicar una regla similar para *Frequency* y *Monetary*

Finalmente formamos una puntuación en RFM (*scores*) sumándolos (se podrían ponderar según la importancia de *Recency*, *Frequency* y *Monetary Value*)

Esto es lo que hacemos con la función RFM que tenemos en el guión `marketing-models.R`.

```{r RFM, echo=FALSE}
source("marketing-models.R")
BBBClub.holdout <- RFM(BBBClub.holdout)
#head(BBBClub.holdout)
BBBClub.holdout$RFM<-BBBClub.holdout$R +BBBClub.holdout$F+BBBClub.holdout$M
knitr::kable(t(head(BBBClub.holdout)), digits=2, caption = "Descripción de las 6 primeras empresas de los nuevos datos ")
```

## Plots
```{r RFM-plots}
library(ggplot2)
ggplot(BBBClub.holdout, aes(R)) +
  geom_histogram()
ggplot(BBBClub.holdout, aes(F)) +
  geom_histogram()
ggplot(BBBClub.holdout, aes(M)) +
  geom_histogram()
ggplot(BBBClub.holdout, aes(RFM)) +
  geom_histogram()

```

# The linear model

Now we calibrate the model using a Linear regression model. We use the choice data set (historic data)

```{r}
BBBClub.lm <- lm(choice~gender+amount+freq+last+first+child+youth+cook+diy+art, data=BBBClub.choice)

#install.packages("stargazer")
library(stargazer)
stargazer(BBBClub.lm, type="text")
```

Use the calibrated model to predict choice on the  holdout sample (we predict the probability of buying the art book using de model from choice data)

```{r}
BBBClub.holdout$plm <- predict(BBBClub.lm, BBBClub.holdout)
knitr::kable(
  t(head(BBBClub.holdout)), 
  digits=2, 
  caption = "Descripción de los 6 primeras individuos de los nuevos datos")

```

# The logistic model

Use the logistic model with the holdout sample (we predict the probability of buying the art book using de model from choice data using a Logistic regression model)

```{r}

BBBClub.glm<-glm(
  choice~gender+amount+freq+last+first+child+youth+cook+diy+art, 
  family=binomial(), 
  data=BBBClub.choice)
stargazer(BBBClub.glm, type="text" )
BBBClub.holdout$pglm <- predict(
  BBBClub.glm, 
  BBBClub.holdout, 
  type="response")

```

## Test accuracy of prediction

```{r}
confusion.glm(BBBClub.choice, BBBClub.glm)
```

## See the new set of variables

```{r}
knitr::kable(
  t(head(BBBClub.holdout)), 
  digits=2, 
  caption = "Descripción de los 6 primeras individuos de los nuevos datos")
```

# Apply the calibrated models to make marketing decisions:  select the segments that will be the target  of our marketing efforts


## Divide individuals in 10 segments according to quantiles

```{r}
quantile(BBBClub.holdout$RFM, probs=seq(.1,1,.1))
quantile(BBBClub.holdout$plm, probs=seq(.1,1,.1))
```

## Use the models por selecting the target of our marketing efforts

### Number of people that bought the book in the choice sample

```{r}
exit<-sum(BBBClub.holdout$choice)
exit
```

## Using the logistic model

```{r}
library(tidyverse)
BBBClub.holdout.bypglm<-arrange(BBBClub.holdout, desc(pglm))

knitr::kable(
  t(head(BBBClub.holdout.bypglm)), 
  digits=2, 
  caption = "Descripción de los 6 primeras individuos de los nuevos datos")


#Prepare the report for logistic model
source("marketing-models.R")

```

## save quartiles

The `ntile` function creates groups according to a variable. In this case the variable is `BBBClub.holdout.bypglm$pglm` and will form 10 groups. Using the argument `reverse=TRUE` it would be necessary the function `reverse.quartile`.

The function `reverse.quartile` in the `marketing-models.R` script changes the order or quartiles, so that the ones with the highest probability are placed in the first decile, and adds to the data base a new variable call `decil`.


```{r}
BBBClub.holdout.bypglm$quartile <- ntile(BBBClub.holdout.bypglm$pglm, 10) 

BBBClub.holdout.bypglm <- reverse.quartile(BBBClub.holdout.bypglm)
```

## check

```{r}
knitr::kable(t(head(BBBClub.holdout.bypglm)), digits=2, caption = "Descripción de los 6 primeras individuos de los nuevos datos")
```

## Cummulative distribution of success (what would have happened if we had used glm to target segments)

```{r}

BBBClub.holdout.bypglm$choice2<-BBBClub.holdout.bypglm$choice/sum(BBBClub.holdout.bypglm$choice)

BBBClub.holdout.bypglm$acumul<-cumsum(BBBClub.holdout.bypglm$choice2)

knitr::kable(t(head(BBBClub.holdout.bypglm)), digits=2, caption = "Descripción de los 6 primeras individuos de los nuevos datos")
```

## by quartile count the success ratio

```{r}
table1<-BBBClub.holdout.bypglm %>%
  group_by(decil) %>%
  summarize(
    count=n(),
    mean.choice=sum(choice)/204
  )
table1
```

## Add the cumulative distribution 

```{r}

table1$acumul<-cumsum(table1$mean.choice)
table1
```

## Add the cumulative mailing to be sent

```{r}

units.mailed<- c(5000,10000,15000,20000,25000,30000,35000,40000,45000,50000)
#units.mailed
#add them to table
table1$mailed<-units.mailed

```

## Now we add the cost, the market potencial, the units sold, the margin  and profit

```{r}
#Add cost
table1$cost<-table1$mailed*0.65
#Add market potential
market.potential<-50000*(exit/2300)
#Add units sold
table1$sold<-table1$acumul*(market.potential)
#table1
#Add profit
margin<-31.95-1.45*15
#margin
table1$profit<-table1$sold*margin-table1$cost
table1
plot(table1$profit)
```

## Check if we improve the profit seleccting onlya few segments

```{r}
print("Total Income")
sum(table1$profit)
print("Total Costs")
sum(table1$cost)
print("Total Profit")
total.profit1<-sum(table1$profit) - sum(table1$cost)
total.profit1
print("Total relative profit")
rendimiento1<-total.profit1/sum(table1$cost)
rendimiento1
```

## Use linear model for selecting the segments to be the target of marketing efforts

### order the data base for linear model

```{r}
BBBClub.holdout.byplm<-arrange(BBBClub.holdout, desc(plm))
knitr::kable(t(head(BBBClub.holdout.byplm)), digits=2, caption = "Descripción de los 6 primeras individuos de los nuevos datos")


```


### Prepare the report for linear model

```{r}
source("marketing-models.R")

#save quartiles
BBBClub.holdout.byplm$quartile <- ntile(BBBClub.holdout.byplm$plm, 10) 
BBBClub.holdout.byplm <- reverse.quartile(BBBClub.holdout.byplm)
#check
knitr::kable(t(head(BBBClub.holdout.byplm)), digits=2, caption = "Descripción de los 6 primeras individuos de los nuevos datos")

```

### Cummulative probability of success

```{r}
BBBClub.holdout.byplm$choice2<-BBBClub.holdout.byplm$choice/sum(BBBClub.holdout.byplm$choice)
BBBClub.holdout.byplm$acumul<-cumsum(BBBClub.holdout.byplm$choice2)

```

### Prepare the table for seleccting the segments

```{r}
#by quartile count the exit ratio

table.lm<-BBBClub.holdout.byplm %>%
  group_by(decil) %>%
  summarize(
    count=n(),
    mean.choice=sum(choice)/204
  )
table.lm
```

### Add the cumulative distribution

```{r}
 
table.lm$acumul<-cumsum(table.lm$mean.choice)
table.lm
```

### Add the cumulative mailing to be sent and add it to the table

```{r}

units.mailed<- c(5000,10000,15000,20000,25000,30000,35000,40000,45000,50000)
units.mailed
#add them to table
table.lm$mailed<-units.mailed
```

### Add cost, market potential, units sold, margin and profit tothe table

```{r}
table.lm$cost<-table.lm$mailed*0.65
#Add market potential
market.potential<-50000*(exit/2300)
#Add units sold
table.lm$sold<-table.lm$acumul*(market.potential)
table.lm
#Add profit
margin<-31.95-1.45*15
margin
table.lm$profit<-table.lm$sold*margin-table.lm$cost
table.lm
plot(table.lm$profit)
```

### Check if we improve the profit seleccting onlya few segments

```{r}
print("Total Income")
sum(table.lm$profit)
print("Total cost")
sum(table.lm$cost)
total.profit.lm<-sum(table.lm$profit) - sum(table.lm$cost)
print("Total profit")
total.profit.lm
rendimiento.lm<-total.profit.lm/sum(table.lm$cost)
print("Relative profit")
rendimiento.lm
```


## RFM model report 

### order the data base for RFM model

```{r}
BBBClub.holdout.byRFM<-arrange(BBBClub.holdout, desc(RFM))

knitr::kable(t(head(BBBClub.holdout.byRFM)), digits=2, caption = "Descripción de los 6 primeras individuos de los nuevos datos")
#Prepare the report for logistic model
source("marketing-models.R")
```

### Devide data according to quartiles

```{r}
#save quartiles
BBBClub.holdout.byRFM$quartile <- ntile(BBBClub.holdout.byRFM$RFM, 10) 
BBBClub.holdout.byRFM <- reverse.quartile(BBBClub.holdout.byRFM)

#check
knitr::kable(t(head(BBBClub.holdout.byRFM)), digits=2, caption = "Descripción de los 6 primeras individuos de los nuevos datos")
```

### Cummulative distribution of success

```{r}
BBBClub.holdout.byRFM$choice2<-BBBClub.holdout.byRFM$choice/sum(BBBClub.holdout.byRFM$choice)
BBBClub.holdout.byRFM$acumul<-cumsum(BBBClub.holdout.byRFM$choice2)

```

### Prepare the summary table

```{r}
#by quartile count the exit ratio

table.RFM<-BBBClub.holdout.byRFM %>%
  group_by(decil) %>%
  summarize(
    count=n(),
    mean.choice=sum(choice)/204,
    mean.RFM=mean(RFM)
  )
table.RFM
```

### Add the cumulative distribution 

```{r}
#Add the cumulative distribution 
table.RFM$acumul<-cumsum(table.RFM$mean.choice)
table.RFM
```


```{r}
#Add the cumulative mailing to be sent
units.mailed<- c(5000,10000,15000,20000,25000,30000,35000,40000,45000,50000)
#units.mailed
#add them to table
table.RFM$mailed<-units.mailed
table.RFM
```

### Add cost, units sold, margin, and profit

```{r}
#Add cost
table.RFM$cost<-table.RFM$mailed*0.65
#Add market potential
market.potential<-50000*(exit/2300)
#Add units sold
table.RFM$sold<-table.RFM$acumul*(market.potential)
table.RFM
#Add profit
margin<-31.95-1.45*15
margin
table.RFM$profit<-table.RFM$sold*margin-table.RFM$cost
table.RFM
plot(table.RFM$profit)
```

# Compare results

```{r}
print("Total income:")
sum(table.RFM$profit)
print("Total cost:")
sum(table.RFM$cost)
print("Total profit:")
total.profit.RFM<-sum(table.RFM$profit) - sum(table.RFM$cost)
total.profit.RFM
print("Total relative profit:")
rendimiento.RFM<-total.profit.RFM/sum(table.RFM$cost)
rendimiento.RFM
```

# Compare findings

## The glm model

```{r}
knitr::kable(table1, digits=2, caption = "Logistic model")
knitr::kable(table.lm, digits=2, caption = "Linear model")
knitr::kable(table.RFM, digits=2, caption = "RFM model")
```



